The Shape Matching Element Method for Meshless Animation of NURBs Models

### 2021-01-17 paper/sections/methods.tex
Filled out most of shape matching part of methods

\section{Methods}
Brief overview. We have some control points. We have UV samples on the NURBS parametric domain. We have a NURBS jacobian UV to points on surface in R3.
Given a volumetric model composed of one or more NURBS surfaces, the surface of each surface is reconstructed using a set of control points $\textbf{p}$ and their associated weights $w$. Following D-NURBS \cite{10.1145/176579.176580}, we take our degrees of freedom (DOF) to be control points of the surfaces and produce displacements directly on these control points at each simulation step. We note that we used a simplified version of D-NURBS in which the weights are held constant throughout the simulation and find this not to be limiting \myworries{(feel like we'll need better justification than saying just this?)}. We represent the degrees of freedom for the $i$-th NURBS surface with $m$ control points as $\mathbf{q}_i = \left[ \mathbf{p}_1^T, \dots, \mathbf{p}_m^T \right]^T \in \mathbb{R}^{3m}$. From this definition, we write the DOF for the entire model as $\mathbf{q} = \left[ \mathbf{q}_i^T, \dots, \mathbf{q}_n^T \right]^T$ where $n$ is the number of NURBS surfaces for the model.

@@

\begin{equation}
\label{eqn:jacobian_diagonal}
    N_{i,j}(u,v)
    = \frac{w_{i,j} B_{i,k}(u)B_{j,l}(v)}{\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
    \text{.}
\end{equation}
Thus the NURBS jacobian for a single $(u,v)$ pair is written as
\begin{equation}
\label{eqn:uv_jacobian}
    \mathbf{J}(u,v)
    \left[ \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{1,1}} \dots
           \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}} \dots 
           \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{n,m}}
    \right]
    \text{.}
\end{equation}

@@

Let's say for a NURBS surface on the model the DOF column vector is $\mathbf{q}_i$, which is the subset of control points in the generalized coordinates for the $i$-th surface. For a given surface with $m$ pairs of $\mathbf{u}=(u,v)$ coordinates sampled in parametric space, we write the jacobian for the $i$-th surface as \myworries{(should I use something like $\hat{\mathbf{J}}$ instead?)}
\begin{equation}
\label{eqn:surface_jacobian}
    \mathbf{J}_i =
    \left[ \mathbf{J}(\mathbf{u})_1^T \dots \mathbf{J}(\mathbf{u})_m^T \right]^T
    \text{.}
\end{equation}
Then if $\mathbf{x}_i$ is the set of $m$ world space positions for the $i$-th surface, we may write $\mathbf{x}_i = \mathbf{J}_i \mathbf{q}_i$. Finally, given the full set of generalized coordinates $\mathbf{q} = \left[ \mathbf{q}_i^T, \dots, \mathbf{q}_n^T \right]^T$ with $n$ surfaces, we may write jacobian $\mathbf{J}$ for the entire model as a block diagonal matrix:
\begin{equation}
\mathbf{J} = \left( \begin{array}{ccccc}
\mathbf{J}_1 &  &  &  &  \\
 & \ddots &  &  &  \\
 &  & \mathbf{J}_i & &  \\
 &  &  & \ddots &  \\
 &  &  &  & \mathbf{J}_n \\
\end{array} \right)
\end{equation}
With this equation, the full set of world positions of the NURBS model may be written as $\mathbf{x} = \mathbf{J}\mathbf{q} \in \mathbb{R}^{3N}$ where $N$ is the total number of $(u,v)$ coordinates.

\section{Shape Matching}
\section{Blending Polynomials}
\section{Raycasting Quadrature}
\section{Dynamic Simulation}
\subsection{Generalized Inertia}
\subsection{Generalized Forces}
\subsection{Equations of Motion}
%This is most notable in cases where the degree of the polynomial on a NURBS is higher than the degree of the B-spline surface.

>>>>

\section{Methods}
Given a volumetric model composed of one or more NURBS surfaces, the surface of each surface is reconstructed using a set of control points $\textbf{p}$ and their associated weights $w$. Following D-NURBS \cite{10.1145/176579.176580}, we take our degrees of freedom (DOF) to be control points of the surfaces and produce displacements directly on these control points at each simulation step. We note that we used a simplified version of D-NURBS in which the weights are held constant throughout the simulation and find this not to be limiting \myworries{(feel like we'll need better justification than saying just this?)}. We represent the degrees of freedom for the $i$-th NURBS surface with $m$ control points as $\mathbf{q}_i = \left[ \mathbf{p}_1^T, \dots, \mathbf{p}_m^T \right]^T \in \mathbb{R}^{3m}$. From this definition, we write the DOF for the entire model as $\mathbf{q} = \left[ \mathbf{q}_i^T, \dots, \mathbf{q}_n^T \right]^T$ where $n$ is the number of NURBS surfaces for the model.

@@

\begin{equation}
\label{eqn:jacobian_diagonal}
    N_{i,j}(u,v)
    = \frac{w_{i,j} B_{i,k}(u)B_{j,l}(v)}{\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
    \text{.}
\end{equation}
Thus the NURBS jacobian for a single $(u,v)$ pair is written as
\begin{equation}
\label{eqn:uv_jacobian}
    \mathbf{J}(u,v) =
    \left[ \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{1,1}} \dots
           \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}} \dots 
           \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{n,m}}
    \right]
    \text{.}
\end{equation}

@@

Let's say for a NURBS surface on the model the DOF column vector is $\mathbf{q}_i$, which is the subset of control points in the generalized coordinates for the $i$-th surface. For a given surface with $m$ pairs of $\mathbf{u}=(u,v)$ coordinates sampled in parametric space, we write the jacobian for the $i$-th surface as \myworries{(should I use something like $\hat{\mathbf{J}}$ instead?)}
\begin{equation}
\label{eqn:surface_jacobian}
    \mathbf{J}_i =
    \left[ \mathbf{J}(\mathbf{u})_1^T \dots \mathbf{J}(\mathbf{u})_m^T \right]^T
    \text{.}
\end{equation}
Then if $\mathbf{x}_i$ is the set of $m$ world space positions for the $i$-th surface, we may write $\mathbf{x}_i = \mathbf{J}_i \mathbf{q}_i$. Finally, given the full set of generalized coordinates $\mathbf{q} = \left[ \mathbf{q}_i^T, \dots, \mathbf{q}_n^T \right]^T$ with $n$ surfaces, we may write jacobian $\mathbf{J}$ for the entire model as a block diagonal matrix:
\begin{equation}
\mathbf{J} = \left[ \begin{array}{ccccc}
\mathbf{J}_1 &  &  &  &  \\
 & \ddots &  &  &  \\
 &  & \mathbf{J}_i & &  \\
 &  &  & \ddots &  \\
 &  &  &  & \mathbf{J}_n \\
\end{array} \right]
\text{.}
\end{equation}
With this equation, the full set of world positions of the NURBS model may be written as $\mathbf{x} = \mathbf{J}\mathbf{q} \in \mathbb{R}^{3N}$ where $N$ is the total number of $(u,v)$ coordinates.

\section{Shape Matching}
The key feature of our method is its unique deformation map $\phi(\mathbf{X}) = \mathbf{x}$ mapping some arbitrary undeformed material point $\mathbf{X}$ to its corresponding deformed position in space $\mathbf{x}$. We develop this similarly to the VEM approach in that we say the function we intend to reproduce may be some highly nonlinear one, but it may be approximated accurately by a polynomial. Thus we seek some $\phi$ function where  $\phi^* \approx \phi$ where $\phi^*$ is the true solution. In $1$-D this means we wish to find a function of the form $\phi(X) = c_1 + c_2X + c_3X^2 + \dots + c_nX^{n-1}$ where each $c_i$ is an unknown polynomial coefficient and the $X^i$ terms are elements of the polynomial basis.

\subsection{The Shape Matching Element}
In typical FEM in graphics, we make use of tetrahedral or triangular elements with a piecewise-linear function space. In such a case our functions may be evaluated as a weighted combination of the nodal values of an element: $f(\mathbf x)=\sum_{i=0}^n f_i \phi_i(\mathbf{x})$ where $n$ is the number of nodal values and $\phi_i$ is the i-th barycentric coordinates. Thus in this case we have the simple deformation map $\mathbf{x(X)} = \sum_{i=0}^n \mathbf{x}_i \phi_i(\mathbf{X})$. However, our method contrasts this typical approach in that our elements are more general and take the form of any arbitrary boundary primitive, such as NURBS, and also that the function space in non-linear.

In our algorithm, a single shape matching element is represented by a NURBS surface, a center of mass, and polynomial describing its current deformation. We construct these polynomials via a shape matching strategy. With the addition of the center of mass, we describe the shapes deformation about this point and thus the monomial basis in the case of a linear polynomial is $\left[ \mathbf 1^T (\mathbf{X - \bar{X}})^T \right]$ where $\mathbf 1^T$ is a constant $1$ column vector and $\mathbf{\bar{X}}$ is the center of mass for the element. To simplify the following formulas, we collect the non-constant terms of the monomial basis into a single vector $\mathbf{m}(\mathbf{X}-\mathbf{\bar{X}})$. For example if $\mathbf{X} = (X,Y,Z)$, $\mathbf{\bar{X}} = (\bar X, \bar Y, \bar Z)$, and $\mathbf{q}=\mathbf{X}-\mathbf{\bar{X}}$ with a quadratic polynomial we would have
\begin{equation}
m(\mathbf{q)} = (
					q_1, q_2, q_3,
                   	q_1^2, q_2^2, q_3^3,
                   	q_1 q_2, q_2 q_3
                ) \in \mathbb{R}^k
                \text{,}
\end{equation}
where $k$ is the size of the monomial basis. Next we arrange this basis into a matrix of the form
\begin{equation}
\mathbf{M}(\mathbf{q}) = \left[ \begin{array}{ccc}
\mathbf{m}(\mathbf{q}) &  &  \\
 & \mathbf{m}(\mathbf{q})  &  \\
 &  & \mathbf{m}(\mathbf{q})  \\
\end{array} \right] \in \mathbb{R}^{3x3k}
\text{.}
\end{equation}
With this definition of the monomial basis, we may reconstruct the deformed position $\mathbf{x}$ for some material point $\mathbf{X}$ with
\begin{equation}
\mathbf{x}(\mathbf{X}) = \left[\mathbf{M}(\mathbf{X-\bar{X}}) \quad \mathbf{I} \right]
\left[ \begin{array}{c}
\mathbf{c} \\
\mathbf{t} \\
\end{array} \right]
\text{,}
\end{equation}
where $\mathbf{c} \in \mathbb{R}^{3k}$ is the set of polynomial coefficients for the non-constant terms, $\mathbf{t} \in \mathbb{R}^3$ is the constant term coefficients, and $\mathbf{I} \in \mathbb{R}^{3x3}$ is an identity matrix. The reason for separating out the constant term becomes more evident when we examine case where we have multiple elements.

The remaining step in defining a single element is to describe how to solve for the polynomial coefficients. For a single element, we have $n$ material points $(\mathbf{X}_1 \dots \mathbf{X}_n)$ and their corresponding deformed positions $(\mathbf{x}^*_1 \dots \mathbf{x}^*_n)$. As these elements are defined on a boundary, these are both known throughout the simulation. Therefore we may form a fitting problem to solve for the polynomial coefficients that minimize the error $\sum_i^n ||\mathbf{x}_i^* - \mathbf{X}_i ||^2$.  We minimize this error in a least squares sense where we first form the system
\begin{equation}
\left[\begin{array}{c}
\mathbf{M}(\mathbf{X_1-\bar{X}}) \quad \mathbf{I} \\
\vdots \\
\mathbf{M}(\mathbf{X_i-\bar{X}}) \quad \mathbf{I} \\
\vdots \\
\mathbf{M}(\mathbf{X_n-\bar{X}}) \quad \mathbf{I} \\
\end{array} \right]
\left[ \begin{array}{c}
\mathbf{c} \\
\mathbf{t} \\
\end{array} \right] = 
\left[ \begin{array}{c}
\mathbf{x}_1^* \\
\vdots \\
\mathbf{x}_i^* \\
\vdots \\
\mathbf{x}_n^*
\end{array} \right]
\text{,}
\end{equation}
and then in the more compact form
\begin{equation}
\left[
\mathbf{A} \quad \mathbf{S}
\right]
\left[ \begin{array}{c}
\mathbf{c} \\
\mathbf{t} \\
\end{array} \right] = \mathbf{b}
\text{,}
\end{equation}
where $\mathbf{A}$ is the block containing the monomial matrices of the system, and $\mathbf{S}$ is the block of identity matrices. Next, to reduce the degrees of freedom of this system we let the deformed center of mass (the constant term coefficients), $t$, be equal to the mean of the deformed positions influenced by this center of mass. In the case of one element we have $t = \frac{1}{n}\sum_i^n x_i^*$. We describe the details on center of mass selection in section \myworries{insert here}. In matrix form we have
\begin{equation}
t = 
\underbrace{
\left[\begin{array}{ccccccc}
\frac{1}{n} & & \cdots \frac{1}{n} & & \\
& \frac{1}{n} & \cdots & \frac{1}{n} & \\
& & \frac{1}{n} \cdots & & \frac{1}{n} \\
\end{array} \right]}_{T} 
\mathbf{b}
\text{.}
\end{equation} (\myworries{spacing on this looks bad....}). Finally, this leaves us with the system
\begin{equation}
\mathbf{A}\mathbf{c}=(\mathbf{I}-\mathbf{ST})\mathbf{b}
\text{.}
\end{equation}
Like in the original Shape Matching \cite{10.1145/1073204.1073216}, we solve for the coefficients by minimizing the error in a least squares sense, resulting in
\begin{equation}
\begin{split}
\mathbf{c} & = (\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T(\mathbf{I-ST})\mathbf{b} \\
           & = \mathbf{L}\mathbf{b}
\end{split}
\text{.}
\end{equation}
This $\mathbf{L}$ matrix thus gives us our file ingredient in the element, coefficients defining the polynomial. We emphasize that $\mathbf{L}$ is constant throughout the simulation, therefore as $\mathbf{b}$ varies in the simulation, $\mathbf{Lb}$ gives us a new set of polynomial coefficients as $\mathbf{b}$ changes.

\subsection{Shape Matching with Multiple Elements}
\subsection{Hierarchical fitting}
A problem may emerge in the shape matching process where the shape matching system may be rank deficient.

\section{Blending Polynomials}
\section{Raycasting Quadrature}
\section{Dynamic Simulation}
\subsection{Generalized Inertia}
\subsection{Generalized Forces}
\subsection{Equations of Motion}
%This is most notable in cases where the degree of the polynomial on a NURBS is higher than the degree of the B-spline surface.
